A load balancing scheme for massively multiplayer online games \thanks{This work was supported by the National Research Council (CNPq) and by the Coordination of Improvement of Higher Education (CAPES), both brazilian research funding agencies.}

### 2009-04-10 trunk/textos/dissertation/MTAP_LaTeX/pub/tex/latex/svjour3/global/balancing.tex

As we said in section \ref{sec:related}, we also compare our algorithms to the one proposed by Ahmed and Shirmohammadi \cite{ahmed2008mol}. However, it is very important to observe that some changes have been made before simulating. First, the authors divide the virtual environment in hexagonal cells, and we simulate their algorithm on a grid-like division, with square cells. Moreover, as we consider an heterogeneous system, each server $S_i$ presents its own message rate threshold, $T_i^m$.
%
Nonetheless, we keep what we consider as the core of their approach, which is the selection of the smallest cell cluster managed by the overloaded server, followed by the selection of the cell of that cluster which has the lowest interaction with other cells of the same cluster and, finally, the transference of the selected cell to the least loaded server.

@@

In the graph of Figure \ref{gra:baloverall}, we can see that all the algorithms proposed met the proportionality criterion in load balancing. However, some of them introduced more overhead than the others. The reason for this is the fragmented regions. The lower the number of fragments that have the regions, the lower is this overhead because there are fewer boundaries between regions.

@@

@@

However, the introduction of overhead is not the only criterion considered. In figure \ref{gra:balmigxtb} is shown how was the migration of users between servers throughout the simulated game session, for each algorithm used in phase 2. The value of migration has been divided in two: \emph{walk migration}, which occurs when a player exchange server because he moved his avatar from one region to another, and \emph{still migration}, which occur when a player exchanges servers without having moved. This kind of migration occurs because the cell where his avatar was has been transferred to another region as a result of a rebalancing of the load of the game. Walk migrations are more likely to happen when the regions are not contiguous.

@@

We can see that \ggp\ is the one which has fewer walk migrations, precisely because their regions are contiguous. However, the fact of not trying to minimize the transfer of cells -- the \ggp\ main objective is to minimize the overhead -- the number of migrations is still the largest of all its variations, lower only than \bfa, which was also the worst algorithm on the user migration criterion. The strategy of \ggpk\, and more strongly in the \ggpf\, to maintain the maximum possible cells during rebalancing, has as a result the lowest numbers of migration of players of all the simulated algorithms.

@@

Even with the servers providing the largest variation in the rate of use of its resources, the \ggp\ was the one with the least overall real weight (which includes overhead: $W_{total} + Overhead$) of all proposed algorithms. Its variant \ggpf\ had the least migrations of users, in exchange for an increase in overhead by 48\%, on the average, compared to the original version of the algorithm. Which of these two is better depends on the specific game. In a real-time game , users contantly migrating between servers can introduce delay and hinder the interaction between players and the \ggpf, as shown, presented a total number of migrations 22\% less than \ggp. Furthermore, it may not be possible to use some kind of ``graceful degradation'' to reduce the quality of the game, needing to save up the servers' resources. In this case, the \ggp\ would be the best option.

>>>>

As we said in section \ref{sec:related}, we also compare our algorithms to the one proposed by Ahmed and Shirmohammadi \cite{ahmed2008mol}. However, it is very important to observe that some changes have been made before simulating. First, the authors divide the virtual environment in hexagonal cells, and we simulate their algorithm on a grid-like division, with square cells. Moreover, as we consider an heterogeneous system, each server $S_i$ presents its own message rate threshold, $T_i^m$.
%
Nonetheless, we keep what we consider the core of their approach, which is the selection of the smallest cell cluster managed by the overloaded server, followed by the selection of the cell of that cluster which has the lowest interaction with other cells of the same cluster and, finally, the transference of the selected cell to the least loaded server.

@@

In Figure \ref{gra:baloverall}, we can see that all the algorithms simulated met the proportionality criterion in load balancing. However, some of them introduced more overhead than the others. The reason for this is the fragmented regions. The lower the number of fragments the regions have, the lower is this overhead because there are fewer boundaries between regions.

@@

Ahmed's algorithm, in turn, presented the second highest average overhead. The most probable reason for this is the choice of the least loaded server to receive a microcell being transferred from an overloaded server. Though this is an obvious choice in terms of load balance, it does not consider that the chosen server may have no connection to that microcell, creating a fragmented region and, thus, increasing the overhead on the system.

@@

However, the introduction of overhead is not the only criterion considered. Fig. \ref{gra:balmigxtb} shows how was the migration of users between servers throughout the simulated game session, for each algorithm. The value of migration has been divided in two: \emph{walking migration}, which occurs when a player migrates to a new server because he moved his avatar from one region to another, and \emph{still migration}, which occur when a player exchanges servers without having moved. This kind of migration occurs because the cell where his avatar was has been transferred to another region as a result of a rebalancing of the load of the game. Walking migrations are more likely to happen when the regions are not contiguous.

@@

We can see that \ggp\ is the one which has fewer walk migrations, precisely because their regions are contiguous. However, the fact of not trying to minimize the transfer of cells -- the \ggp\ main objective is to minimize the overhead -- the number of migrations is still the largest of all its variations, lower only than \bfa, which was also the worst algorithm on the user migration criterion. The strategy of \ggpk, and more strongly in the \ggpf, to maintain the maximum possible number of cells when rebalancing, has as a result the lowest numbers of migration of players of all the proposed algorithms.

The algorithm proposed by Ahmed had even less migrations than \ggpf. Though the number of walk migrations is the second highest -- as expected from a more fragmented partitioning --, the number of still migrations is almost negligible. We believe that a clever choice of cells to transfer led to this result. Cells which interact less with other cells will, probably, contain less avatars and, then, trasferring them will not cause many player migrations.

@@

Even with the servers providing the largest variation in the rate of use of its resources, the \ggp\ was the one with the least overall real weight (which includes overhead: $W_{total} + OverHead$) of all proposed algorithms. Its variant \ggpf\ had the least migrations of users, in exchange for an increase in overhead by 48\%, on the average, compared to the original version of the algorithm. Which of these two is better depends on the specific game. In a real-time game , users contantly migrating between servers can introduce delay and hinder the interaction between players and the \ggpf, as shown, presented a total number of migrations 22\% less than \ggp. Furthermore, it may not be possible to use some kind of ``graceful degradation'' to reduce the quality of the game, needing to save up the servers' resources. In this case, the \ggp\ would be the best option.
